### 2020-06-22 manuscript.tex
edits 18.6.20
shingii

We collected 90 samples, 56 from herbarium specimens and 34 from silica-dried tissue. These samples represented five of seven named species of \textit{Ceratopteris}, and one sample of \textit{Acrostichum aureum} from Australia for use as an outgroup. Included in our silica-dried specimens was a sample of the common lab strain of \textit{C. richardii} (\textit{Hnn}). The two missing species of \textit{Ceratopteris} were cryptic species of \textit{C. thalictroides} (L.) Brongn. \autocite{Masuyama2010}: \textit{C. oblongiloba} Masuyama \& Watano, known from Southeast Asia; and \textit{C. froesii} Brade, endemic to Brazil. Both have relatively narrow ranges and few herbarium species, especially \textit{C. froesii}. We were unable to obtain any material from herbarium specimens labeled as either of these species, and none of the plants collected in the field keyed out to \textit{C. oblongiloba} or \textit{C. froesii}.

@@

Two named cryptic species of \textit{C. thalictroides} were not included in our analysis: \textit{C. oblongiloba} and \textit{C. froseii}. We were not able to request type material, we had to rely on our herbarium sampling of their ranges to attempt to include these species. \textit{Ceratopteris oblongiloba} is found throughout Cambodia, Indonesia (Sumatra and Java), the Philippines (Luzon), and Thailand (Malay) \autocite{Masuyama2010}. Our only specimen from this range is from the Philippines, does not key to \textit{C. oblongiloba}, and aligns entirely with Old World \textit{C. thalictroides}. This either indicates that we did not unknowingly sample \textit{C. oblongiloba}, or it has a lesser degree of genomic distinction than the other cryptic species, \textit{C. gaudichaudii}. \textit{Ceratopteris froesii} is a Brazilian endemic characterized by very small fronds: about 4 cm or less for fertile leaves \autocite{Masuyama2010}. We only had four accessions from Brazil, and none were small plants. Inclusion of this unique species in future studies may help us further understand New World diversity of \textit{Ceratopteris}.

@@

Supplementary data 1: Specimen information

@@

\begin{figure}[H]
\centering
\includegraphics[scale=0.45]{map2.pdf}
\caption{Collection localities for samples of \textit{Ceratopteris} included in the present study. These collections roughly indicate the range for each species. Not all accessions are included on this map because many herbarium specimens did not have GPS coordinates, or several samples were from nearly identical locations. Localities of \textit{C. richardii} are those of sampled individuals, but were ultimately not included in our analyses.}
\label{map}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=1]{g2p.pdf}
\caption{Ploidy estimates for a subset of individuals. The X axis displays the estimated allele ratios (1:1, 2:1, 3:1), and the Y axis shows the Bayesian posterior proportion of allelic ratios; 95\% credible intervals are shown on each bar. The allelic ratio with the highest posterior estimate was used to assign ploidy to an individual. If the credible intervals overlapped, ploidy was assigned as ambiguous. \textit{Ceratopteris cornuta} (green) and \textit{C. pteridoides} (purple) are both known diploids (\textit{n} = 39; \cite{Hickok1977}). All specimens of these species had a 3:1 allele ratio, indicating potential homoploid hybridization, or genomic restructuring within diploids to yield such unbalanced allele ratios. \textit{C. gaudichaudii} and \textit{C. thalictroides} are known tetraploids (\textit{n} = 77, 78; \cite{Masuyama2010}). \textit{Ceratopteris thalictroides} 1 (grey) had two distinct patterns of allele ratios, a 1:1 (autopolyploid) group and a mixed to ambigious group. The former is entirely from the Old World, whereas the latter group contains individuals from the Old and New Worlds. \textit{Ceratopteris thalictroides} 2 (yellow) is only found in the New World, and all individuals had mixed 1:1 and 3:1 (allopolyploid) allele ratios. See Supplementary data 2 for sample IDs and all ploidy estimates.}
\label{g2p}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=0.7]{k2-6.pdf}
\caption{{\small{STRUCTURE}} plots for \textit{K} = 2 - 6. Each bar represents an individual, and each color is representative of a single genomic population. A single bar with multiple colored segments indicates ancestral admixture from \textit{K} source populations. \textit{K} = 5 was determined to be the most biologically informative. At \textit{K} = 5, all named species of \textit{Ceratopteris} fall into distinct populations. \textit{Ceratopteris thalictroides} is split into three groups: Old World (blue), New World (yellow), and individuals grouping with \textit{C. gaudichaudii} (purple). \textit{Ceratopteris cornuta} is the only species with evidence of ongoing or recent admixture, most evident with \textit{C. pteridoides}. At \textit{K} = 6, the yellow bars may be indicative of \textit{C. richardii}, but we not have a sample here to compare it to; see Supplementary data 2.}
\label{structure}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale=0.65]{tetrad_new_colors.pdf}
\caption{Phylogeny of \textit{Ceratopteris} generated via quartet methods. The tree is rooted with the sister genus to \textit{Ceratopteris}, \textit{Acrostichum}. \textit{Ceratopteris pteridoides} is the most basal clade of the genus, and a clade of \textit{C. cornuta} is the next to diverge. This clade might actually be a sampling of \textit{C. richardii}; however, future sampling is needed to confirm this. Several individuals originally identified as \textit{C. thalictroides} grouped with \textit{C. gaudichaudii}, indicating that this cryptic species is indeed prolific. The othe individuals of \textit{Ceratopteris thalictroides sensu latu} breaks into two groups here: an Old World clade (1) and a New World clade (2). A clade of \textit{C. cornuta} is sister to Old World \textit{C. thalictroides} 1, which supports it as a potential progenitor of the latter species \autocite{Adjie2007}.}
\label{phy}
\end{figure}

>>>>

We collected 90 samples, 56 from herbarium specimens and 34 from silica-dried tissue. These samples represented five of seven named species of \textit{Ceratopteris}, and one sample of \textit{Acrostichum aureum} from Australia for use as an outgroup. Included in our silica-dried specimens was a sample of the common lab strain of \textit{C. richardii} (\textit{Hnn}). The three missing species of \textit{Ceratopteris} were cryptic species of \textit{C. thalictroides} (L.) Brongn. \autocite{Masuyama2010}: \textit{C. oblongiloba} Masuyama \& Watano, known from Southeast Asia; \textit{C. froesii} Brade, endemic to Brazil; and \textit{C. shingii} Y. H. Yan & R. Zhang, endemic to Hainan Provence, China. These species have relatively narrow ranges and few herbarium species, especially \textit{C. froesii}. We were unable to obtain any material from herbarium specimens labeled as either of these species, and none of the plants collected in the field keyed out to \textit{C. oblongiloba} or \textit{C. froesii}. \textit{Ceratopteris shingii} was described as this manuscript was nearing completion, so no material could be included.

@@

Three named cryptic species of \textit{C. thalictroides} were not included in our analysis: \textit{C. oblongiloba}, \textit{C. froesii}, and \textit{C. shingii}. We were not able to request type material, we had to rely on our herbarium sampling of their ranges to attempt to include \textit{C. oblongiloba} and \textit{C. froesii}. \textit{Ceratopteris shingii} was described as this manuscript was nearing completion.

\textit{Ceratopteris oblongiloba} is found throughout Cambodia, Indonesia (Sumatra and Java), the Philippines (Luzon), and Thailand (Malay) \autocite{Masuyama2010}. Our only specimen from this range is from the Philippines, does not key to \textit{C. oblongiloba}, and aligns entirely with Old World \textit{C. thalictroides}. This either indicates that we did not unknowingly sample \textit{C. oblongiloba}, or it has a lesser degree of genomic distinction than the other cryptic species, \textit{C. gaudichaudii}. \textit{Ceratopteris froesii} is a Brazilian endemic characterized by very small fronds: about 4 cm or less for fertile leaves \autocite{Masuyama2010}. We only had four accessions from Brazil, and none were small plants. Inclusion of this unique species in future studies may help us further understand New World diversity of \textit{Ceratopteris}. \textit{Ceratopteris shingii} was recently described by \textcite{Zhang2020}, endemic to Hainan Province, China. It is the only known species in the genus to have creeping rhizomes, and is estimated to be sister to all other species of \textit{Ceratopteris} \autocite{Zhang2020}.

@@

Supplementary data 1: Table A1, Specimen information

@@

\begin{figure}[H]
\centering
%\includegraphics[scale=0.45]{map2.pdf}
\caption{Collection localities for samples of \textit{Ceratopteris} included in the present study. These collections roughly indicate the range for each species. Not all accessions are included on this map because many herbarium specimens did not have GPS coordinates, or several samples were from nearly identical locations. Localities of \textit{C. richardii} are those of sampled individuals, but were ultimately not included in our analyses.}
\label{map}
\end{figure}

\begin{figure}[H]
\centering
%\includegraphics[scale=1]{g2p.pdf}
\caption{Ploidy estimates for a subset of individuals. The X axis displays the estimated allele ratios (1:1, 2:1, 3:1), and the Y axis shows the Bayesian posterior proportion of allelic ratios; 95\% credible intervals are shown on each bar. The allelic ratio with the highest posterior estimate was used to assign ploidy to an individual. If the credible intervals overlapped, ploidy was assigned as ambiguous. \textit{Ceratopteris cornuta} (green) and \textit{C. pteridoides} (purple) are both known diploids (\textit{n} = 39; \cite{Hickok1977}). All specimens of these species had a 3:1 allele ratio, indicating potential homoploid hybridization, or genomic restructuring within diploids to yield such unbalanced allele ratios. \textit{C. gaudichaudii} and \textit{C. thalictroides} are known tetraploids (\textit{n} = 77, 78; \cite{Masuyama2010}). \textit{Ceratopteris thalictroides} 1 (grey) had two distinct patterns of allele ratios, a 1:1 (autopolyploid) group and a mixed to ambigious group. The former is entirely from the Old World, whereas the latter group contains individuals from the Old and New Worlds. \textit{Ceratopteris thalictroides} 2 (yellow) is only found in the New World, and all individuals had mixed 1:1 and 3:1 (allopolyploid) allele ratios. See Supplementary data 2 for all ploidy estimates.}
\label{g2p}
\end{figure}

\begin{figure}[H]
\centering
%\includegraphics[scale=0.7]{k2-6.pdf}
\caption{{\small{STRUCTURE}} plots for \textit{K} = 2 - 6. Each bar represents an individual, and each color is representative of a single genomic population. A single bar with multiple colored segments indicates ancestral admixture from \textit{K} source populations. \textit{K} = 5 was determined to be the most biologically informative. At \textit{K} = 5, all named species of \textit{Ceratopteris} fall into distinct populations. \textit{Ceratopteris thalictroides} is split into three groups: Old World (blue), New World (yellow), and individuals grouping with \textit{C. gaudichaudii} (purple). \textit{Ceratopteris cornuta} is the only species with evidence of ongoing or recent admixture, most evident with \textit{C. pteridoides}. At \textit{K} = 6, the yellow bars may be indicative of \textit{C. richardii}, but we not have a sample here to compare it to; see Supplementary data 2.}
\label{structure}
\end{figure}

\begin{figure}[H]
\centering
%\includegraphics[scale=0.65]{tetrad_new_colors.pdf}
\caption{Phylogeny of \textit{Ceratopteris} generated via quartet methods. The tree is rooted with the sister genus to \textit{Ceratopteris}, \textit{Acrostichum}. \textit{Ceratopteris pteridoides} is the most basal clade of the genus, and a clade of \textit{C. cornuta} is the next to diverge. This clade might actually be a sampling of \textit{C. richardii}; however, future sampling is needed to confirm this. Several individuals originally identified as \textit{C. thalictroides} grouped with \textit{C. gaudichaudii}, indicating that this cryptic species is indeed prolific. The othe individuals of \textit{Ceratopteris thalictroides sensu latu} breaks into two groups here: an Old World clade (1) and a New World clade (2). A clade of \textit{C. cornuta} is sister to Old World \textit{C. thalictroides} 1, which supports it as a potential progenitor of the latter species \autocite{Adjie2007}.}
\label{phy}
\end{figure}
